---
title: "the masteRs (IN PROGRESS) :)"
output: github_document
knit: (function(input, ...) { rmarkdown::render(input, output_file = "README.md", envir = globalenv()) })
---

Data: [Raw Data Archive](https://datagolf.com/raw-data-archive){target="_blank"} via [datagolf.com](https://datagolf.com/){target="_blank"}

___

```{r include = F}
knitr::knit_hooks$set(source = function(x, options) {
    hook.r = function(x, options) {
      fence = "```"
      language = tolower(options$engine)
      if (language == "node") language = "javascript"
      if (!options$highlight) language = "text"
      if (!is.null(options$fold_code)) {
        paste0("\n\n", "<details><summary>View Code</summary>\n", fence, language, "\n", x, fence, "\n\n", "</details>\n")
      } else paste0('\n\n', fence, language, '\n', x, fence,  '\n\n')
    }
    x = knitr:::hilight_source(x, "markdown", options)
    hook.r(paste(c(x, ""), collapse = "\n"), options)
})
```

### Setup

```{r message = F, warning = F, fold_code = T}
library(tidyverse)
library(janitor)
library(tvthemes)
library(plotly)

theme_custom = theme_avatar() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, size = 9, vjust = 2.5, face = "italic"),
        plot.caption = element_text(face = "italic"),
        panel.grid.major = element_line(linewidth = 0.5, colour = "#DFDAD1"),
        panel.grid.minor = element_line(linewidth = 0.5, colour = "#DFDAD1"))

theme_set(theme_custom)
custom_red = "#DA8B8B"
custom_green = "#839F7D"
```

### Data Import

```{r fold_code = T}
# changing player name to be (first + last) instead of (last, first)
df = read_csv("data/masters_data_2021.csv", show_col_types = F) |>
  separate(player_name, into = c("last_name", "first_name"), sep = ",") |>
  mutate(player_name = paste(first_name, last_name)) |>
  select(1:5, 31, 8:30)

paste0("Data dimensions: ", nrow(df), " rows, ", ncol(df), " columns")
```

### Scoring Gaps

Which players had the largest difference between their best and worst scores for the tournament?

```{r fold_code = T}
min_max_rounds = df |>
  group_by(player_name) |>
  summarise(min_round = min(round_score),
            max_round = max(round_score)) |>
  mutate(min_max_diff = max_round - min_round)

min_max_rounds |>
  slice_max(min_max_diff, n = 5) |>
  pivot_longer(c(min_round, max_round), names_to = "min_max", values_to = "score") |>
  ggplot(aes(reorder(player_name, -min_max_diff), score)) +
  geom_col(aes(fill = min_max), position = "dodge", show.legend = F) +
  geom_text(aes(label = score, group = min_max), position = position_dodge2(width = 0.9), vjust = -0.5, size = 3.5) +
  scale_fill_manual(values = c(custom_red, custom_green)) +
  labs(x = NULL, y = "Score",
       title = "Who had the biggest difference in their best and worst rounds?",
       subtitle = "2021 Masters Tournament") +
  theme(axis.text.y = element_blank())
```

**Plot Notes**

- Bernd Wiesberger shot 78 in round 4 and 66 in round 2
- Francesco Molinari shot 81 in round 4 and 69 in round 3
- Carlos Ortiz shot 82 in round 1 and 71 in round 2 (missed cut)
- Joe Long shot 82 in round 1 and 72 in round 2 (missed cut)
- Hudson Swafford shot 83 in round 2 and 73 in round 1 (missed cut)

___

### Great Shots

Who hit the most great shots?

```{r fold_code = T}
# cor_df = df |>
#   filter(!is.na(great_shots) & !is.na(round_score))
# 
# round(cor(cor_df$great_shots, cor_df$round_score), 3)

df |>
  filter(!is.na(great_shots)) |>
  group_by(player_name, fin_text) |>
  summarise(great_shots = sum(great_shots),
            .groups = "drop") |>
  slice_max(great_shots, n = 10) |>
  mutate(p_name = paste0(player_name, " (", fin_text, ")")) |>
  ggplot(aes(reorder(p_name, great_shots), great_shots)) +
  geom_col(fill = custom_green) +
  geom_text(aes(label = great_shots), size = 3.5, hjust = -0.4) +
  coord_flip() +
  labs(x = NULL, y = "Great Shots",
       title = "Who hit the most great shots?",
       subtitle = "2021 Masters Tournament") +
  theme(axis.text.x = element_blank())
```

**Plot Note**: The number of great shots a player hit had a -0.367 correlation coefficient with that player's round score, meaning that as players hit more great shots, their score decreased - just as we'd expect!

___

### Poor Shots

Who hit the most poor shots?

```{r fold_code = T}
# cor_df = df |>
#   filter(!is.na(poor_shots) & !is.na(round_score))
# 
# round(cor(cor_df$poor_shots, cor_df$round_score), 3)

df |>
  filter(!is.na(poor_shots)) |>
  group_by(player_name, fin_text) |>
  summarise(poor_shots = sum(poor_shots),
            .groups = "drop") |>
  slice_max(poor_shots, n = 10) |>
  mutate(p_name = paste0(player_name, " (", fin_text, ")")) |>
  ggplot(aes(reorder(p_name, poor_shots), poor_shots)) +
  geom_col(fill = custom_red) +
  geom_text(aes(label = poor_shots), size = 3.5, hjust = -0.4) +
  coord_flip() +
  labs(x = NULL, y = "Poor Shots",
       title = "Who hit the most poor shots?",
       subtitle = "2021 Masters Tournament") +
  theme(axis.text.x = element_blank())
```

**Plot Note**: The number of poor shots a player hit had a 0.592 correlation coefficient with that player's round score, meaning that as players hit more poor shots, their score increased - this is compared to a -0.367 correlation coefficient between great shots and round score, so the number of poor shots a player hit in a round was more impactful to their score than the number of great shots was.

___

### Great and Poor Shots

What is the relationship between great and poor shots?

```{r fold_code = T}
made_missed_cut = df |>
  filter(!is.na(great_shots) & !is.na(poor_shots)) |>
  count(player_name) |>
  mutate(made_cut = ifelse(n == 4, "Made Cut", "Missed Cut"))

# cor_df = df |>
#   filter(!is.na(great_shots) & !is.na(poor_shots)) |>
#   group_by(player_name) |>
#   summarise(great = sum(great_shots),
#             poor = sum(poor_shots),
#             .groups = "drop")
# 
# round(cor(cor_df$great, cor_df$poor), 3)

df |>
  filter(!is.na(great_shots) & !is.na(poor_shots)) |>
  inner_join(made_missed_cut, by = "player_name") |>
  group_by(player_name, made_cut) |>
  summarise(great = sum(great_shots),
            poor = sum(poor_shots),
            .groups = "drop") |>
  ggplot(aes(great, poor)) +
  geom_point(aes(col = made_cut), size = 4) +
  geom_abline(linetype = "dashed", alpha = 0.5) +
  # geom_line(stat = "smooth", formula = y ~ x, method = "lm", linetype = "dashed", alpha = 0.5) +
  scale_color_manual(values = c(custom_green, custom_red)) +
  labs(x = "Great Shots", y = "Poor Shots", col = NULL,
       title = "Scatterplot of Great and Poor Shots",
       subtitle = "2021 Masters Tournament") +
  theme(legend.position = "right") +
  scale_x_continuous(breaks = seq(0, 50, by = 2)) +
  scale_y_continuous(breaks = seq(0, 50, by = 2))
```

**Plot Note**: Over the course of the tournament, there is a positive correlation of 0.583 between the number of great shots a player hit and the number of poor shots a player hit. In the plot, the dashed line represents the points where the number of great shots is equal to the number of poor shots a player hit - ideally, players want to be below this line, indicating they hit more great shots than poor shots.

___

### Great Shots v Poor Shots

Which players had the biggest difference in great and poor shots, in either direction?

```{r fold_code = T}
n_threshold = 6

df |>
  filter(!is.na(great_shots) & !is.na(poor_shots)) |>
  group_by(player_name, fin_text) |>
  summarise(great = sum(great_shots),
            poor = sum(poor_shots),
            .groups = "drop") |>
  mutate(diff = great - poor) |>
  slice_max(diff, n = n_threshold) |>
  rbind(df |>
  filter(!is.na(great_shots) & !is.na(poor_shots)) |>
  group_by(player_name, fin_text) |>
  summarise(great = sum(great_shots),
            poor = sum(poor_shots),
            .groups = "drop") |>
  mutate(diff = great - poor) |>
  slice_min(diff, n = n_threshold)) |>
  mutate(diff_col = ifelse(diff > 0, "good", "bad"),
         pos_lab = ifelse(diff > 0, diff, ""),
         neg_lab = ifelse(diff < 0, diff, ""),
         p_name = paste0(player_name, " (", fin_text, ")")) |>
  ggplot(aes(reorder(p_name, diff), diff)) +
  geom_col(aes(fill = diff_col), show.legend = F) +
  geom_text(aes(label = pos_lab), size = 3.5, hjust = 1.5) +
  geom_text(aes(label = neg_lab), size = 3.5, hjust = -0.4) +
  coord_flip() +
  scale_fill_manual(values = c(custom_red, custom_green)) +
  labs(x = NULL, y = "Great Shots - Poor Shots",
       title = "Biggest Differences Between Great and Poor Shots",
       subtitle = "2021 Masters Tournament") +
  scale_y_continuous(breaks = seq(-15, 15, by = 3))
```

___

### Impactful Shots by Round

```{r fold_code = T}
df |>
  filter(!is.na(great_shots) & !is.na(poor_shots) & !is.na(round_score)) |>
  inner_join(df |>
  filter(!is.na(great_shots) & !is.na(poor_shots) & !is.na(round_score)) |>
  group_by(player_name, round_num) |>
  summarise(great = sum(great_shots),
            poor = sum(poor_shots),
            impactful = sum(great_shots) + sum(poor_shots),
            .groups = "drop"), by = c("player_name", "round_num")) |>
  mutate(great_prop = great / impactful,
         round_num = paste0("Round ", round_num)) |>
  ggplot(aes(impactful, round_score)) +
  geom_point() +
  geom_smooth(formula = y ~ x, method = "lm", se = F, col = "springgreen4") +
  geom_hline(yintercept = 72, alpha = 0.15) +
  facet_wrap(vars(round_num)) +
  labs(x = "Impactful Shots", y = "Round Score",
       title = "Impactful Shots by Round",
       subtitle = "2021 Masters Tournament") +
  scale_x_continuous(breaks = seq(0, 20, by = 2)) +
  scale_y_continuous(breaks = seq(60, 90, by = 4))
```

**Plot Note**: I am defining an impactful shot as one that has either a great or poor outcome (as determined by the data source). One way of interpreting this may also be considering these as high-risk shots - shots that are risky and can either turn out great or turn out poorly. This may not necessarily be the case for each great or poor shot, but is how I am tracking it with the data at hand. With this interpretation, round three is called moving day for a reason - it is the only round where the number of impactful shots has a negative relationship with that player's score. The relationship is also notably more positive in rounds two and four, which I suspect may be due to course setup or some players hitting higher-risk shots in attempt to make the cut or improve their finishing position.

___

### Scoring by Round and Time of Day

```{r fold_code = T}
df |>
  inner_join(df |>
  distinct(round_num, teetime) |>
  group_by(round_num) |>
  mutate(round_group_num = rank(teetime)), by = c("round_num", "teetime")) |>
  group_by(round_num, teetime) |>
  summarise(score = sum(round_score),
            .groups = "drop") |>
  inner_join(df |>
  group_by(round_num, teetime) |>
  summarise(n_players = n(),
            .groups = "drop"), by = c("round_num", "teetime")) |>
  mutate(avg_score = round(score / n_players, 3),
         round_num = paste0("Round ", round_num)) |>
  ggplot(aes(teetime, avg_score)) +
  geom_line(linewidth = 1.5, col = custom_green) +
  geom_line(stat = "smooth", formula = y ~ x, method = "lm", alpha = 0.5) +
  facet_wrap(vars(round_num)) +
  scale_y_continuous(breaks = seq(60, 90, by = 2)) +
  labs(x = "Tee Time", y = "Group Average Score",
       title = "Scores by Tee Times",
       subtitle = "2021 Masters Tournament")
```

**Plot Note**: I'm not sure there's much to be seen here, as scores appear to vary throughout the day for all four rounds. The only thing I find potentially interesting is that for rounds three and four, the highest scores appear to directly follow the lowest scores for the day. For round two, the same is nearly true, but the lowest scores occur well after the highest scores for the day. I don't make much of it.

___

### Driving Accuracy by Distance

```{r fold_code = T}
# cor_df = df |>
#   filter(!is.na(driving_dist) & !is.na(driving_acc) & round_num == 1)
# 
# round(cor(cor_df$driving_dist, cor_df$driving_acc), 3)
# 
# cor_df = df |>
#   filter(!is.na(driving_dist) & !is.na(driving_acc) & round_num == 2)
# 
# round(cor(cor_df$driving_dist, cor_df$driving_acc), 3)
# 
# cor_df = df |>
#   filter(!is.na(driving_dist) & !is.na(driving_acc) & round_num == 3)
# 
# round(cor(cor_df$driving_dist, cor_df$driving_acc), 3)
# 
# cor_df = df |>
#   filter(!is.na(driving_dist) & !is.na(driving_acc) & round_num == 4)
# 
# round(cor(cor_df$driving_dist, cor_df$driving_acc), 3)

df |>
  filter(!is.na(driving_dist) & !is.na(driving_acc)) |>
  mutate(round_num = paste0("Round ", round_num)) |>
  ggplot(aes(driving_dist, driving_acc)) +
  geom_point() +
  geom_smooth(formula = y ~ x, method = "lm", se = F, col = custom_red) +
  facet_wrap(vars(round_num), scales = "free_x") +
  labs(x = "Driving Distance", y = "Driving Accuracy",
       title = "Driving Accuracy by Distance",
       subtitle = "2021 Masters Tournament") +
  scale_x_continuous(breaks = seq(200, 400, by = 10)) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.1), labels = scales::percent)
```

**Plot Notes**: As we would likely expect, there is a negative relationship between driving distance and driving accuracy. The correlation coefficient between the two by round was pretty consistent over the weekend, going -0.288, -0.261, -0.241, and -0.249 in rounds one through four respectively. This is not necessarily that strong of a correlation, but I find it interesting how consistent the coefficient is over the weekend.

___

```{r}
# cor_df = df |>
#   select(sg_ott, driving_dist, driving_acc) |>
#   na.omit()
# 
# acc_cor = round(cor(cor_df$driving_acc, cor_df$sg_ott), 3)
# dist_cor = round(cor(cor_df$driving_dist, cor_df$sg_ott), 3)

df |>
  select(sg_ott, driving_dist, driving_acc) |>
  na.omit() |>
  pivot_longer(!sg_ott, names_to = "driving_metric", values_to = "value") |>
  mutate(driving_metric = ifelse(driving_metric == "driving_dist", "Driving Distance", "Driving Accuracy")) |>
  ggplot(aes(value, sg_ott)) +
  geom_point(col = custom_green) +
  geom_line(stat = "smooth", formula = y ~ x, method = "lm", linetype = "dashed") +
  geom_hline(yintercept = 0, alpha = 0.25) +
  facet_wrap(vars(driving_metric), scales = "free_x") +
  labs(x = "Metric Value", y = "Strokes Gained Off the Tee",
       title = "Strokes Gained Off the Tee",
       subtitle = "2021 Masters Tournament") +
  scale_y_continuous(breaks = seq(-5, 5, by = 1))
```

**Plot Note**: Both driving accuracy and distance have a positive correlation with strokes gained off the tee, with driving accuracy having a correlation coefficient of 0.534 and driving distance having a correlation of 0.443. Extra distance is nice, but keeping the ball in the fairway looks to have been more important during this weekend at Augusta.

___

### Strokes Gained Metrics

```{r fold_code = T}
# cor_mat = df |>
#   select(contains("sg_")) |>
#   na.omit() |>
#   cor() |>
#   round(3)
# 
# cor_mat[1, 6]
# cor_mat[2, 6]
# cor_mat[3, 6]
# cor_mat[4, 6]
# cor_mat[5, 6]

df |>
  select(contains("sg_")) |>
  na.omit() |>
  pivot_longer(!sg_total, names_to = "sg_metric", values_to = "value") |>
  mutate(sg_metric = case_when(sg_metric == "sg_putt" ~ "Strokes Gained (Putting)",
                               sg_metric == "sg_arg" ~ "Strokes Gained (Around the Green)",
                               sg_metric == "sg_app" ~ "Strokes Gained (Approach)",
                               sg_metric == "sg_ott" ~ "Strokes Gained (Off the Tee)",
                               sg_metric == "sg_t2g" ~ "Strokes Gained (Tee to Green)"),
         sg_metric = factor(sg_metric, levels = c(
           "Strokes Gained (Off the Tee)", "Strokes Gained (Approach)", "Strokes Gained (Around the Green)",
           "Strokes Gained (Tee to Green)", "Strokes Gained (Putting)"))) |>
  ggplot(aes(value, sg_total)) +
  geom_point(aes(col = sg_metric), show.legend = F) +
  geom_line(stat = "smooth", formula = y ~ x, method = "lm", linetype = "dashed") +
  facet_wrap(vars(sg_metric), scales = "free") +
  scale_x_continuous(breaks = seq(-10, 10, by = 2)) +
  scale_y_continuous(breaks = seq(-15, 15, by = 3)) +
  labs(x = "Strokes Gained (X)", y = "Strokes Gained Total",
       title = "Strokes Gained Metrics")
```

**Plot Notes**: As expected, each of the individual strokes gained metrics have a positive correlation with total strokes gained. But which of them (other than tee to green, which is already a combination of more than one strokes gained metric) is most strongly correlation with total strokes gained?

- Off the Tee: 0.400
- Approach: 0.613
- Around the Green: 0.492
- *Tee to Green: 0.832*
- Putting: 0.546

XXX MORE INFO HERE

___

### Strokes Gained Ranks?

```{r}
df |>
  select(player_name, contains("sg_")) |>
  na.omit() |>
  group_by(player_name) |>
  summarise(putt = sum(sg_putt),
            arg = sum(sg_arg),
            app = sum(sg_app),
            ott = sum(sg_ott),
            t2g = sum(sg_t2g),
            total = sum(sg_total)) |>
  mutate(putt_rank = rank(-putt, ties.method = "min"),
         arg_rank = rank(-arg, ties.method = "min"),
         app_rank = rank(-app, ties.method = "min"),
         ott_rank = rank(-ott, ties.method = "min"),
         t2g_rank = rank(-t2g, ties.method = "min"),
         total_rank = rank(-total, ties.method = "min"))
```









