---
title: "the masteRs"
output: github_document
knit: (function(input, ...) { rmarkdown::render(input, output_file = "README.md", envir = globalenv()) })
---

Data: [Raw Data Archive](https://datagolf.com/raw-data-archive){target="_blank"} via [datagolf.com](https://datagolf.com/){target="_blank"}

___

```{r include = F}
knitr::knit_hooks$set(source = function(x, options) {
    hook.r = function(x, options) {
      fence = "```"
      language = tolower(options$engine)
      if (language == "node") language = "javascript"
      if (!options$highlight) language = "text"
      if (!is.null(options$fold_code)) {
        paste0("\n\n", "<details><summary>View Code</summary>\n", fence, language, "\n", x, fence, "\n\n", "</details>\n")
      } else paste0('\n\n', fence, language, '\n', x, fence,  '\n\n')
    }
    x = knitr:::hilight_source(x, "markdown", options)
    hook.r(paste(c(x, ""), collapse = "\n"), options)
})
```

### Setup

```{r message = F, warning = F, fold_code = T}
library(tidyverse)
library(janitor)
library(tvthemes)
library(plotly)

theme_custom = theme_avatar() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, size = 9, vjust = 2.5, face = "italic"),
        plot.caption = element_text(face = "italic"),
        panel.grid.major = element_line(linewidth = 0.5, colour = "#DFDAD1"),
        panel.grid.minor = element_line(linewidth = 0.5, colour = "#DFDAD1"))

theme_set(theme_custom)
custom_red = "#DA8B8B"
custom_green = "#839F7D"
```

### Data Import

```{r fold_code = T}
df = read_csv("data/masters_data_2021.csv", show_col_types = F)
paste0("Data dimensions: ", nrow(df), " rows, ", ncol(df), " columns")
```

### Scoring Gaps

Which players had the largest difference between their best and worst scores for the tournament?

```{r fold_code = T}
min_max_rounds = df |>
  group_by(player_name) |>
  summarise(min_round = min(round_score),
            max_round = max(round_score)) |>
  mutate(min_max_diff = max_round - min_round)

min_max_rounds |>
  slice_max(min_max_diff, n = 5) |>
  pivot_longer(c(min_round, max_round), names_to = "min_max", values_to = "score") |>
  ggplot(aes(reorder(player_name, -min_max_diff), score)) +
  geom_col(aes(fill = min_max), position = "dodge", show.legend = F) +
  geom_text(aes(label = score, group = min_max), position = position_dodge2(width = 0.9), vjust = -0.5, size = 3.5) +
  scale_fill_manual(values = c(custom_red, custom_green)) +
  labs(x = NULL, y = "Score",
       title = "Who had the biggest difference in their best and worst rounds?",
       subtitle = "2021 Masters Tournament") +
  theme(axis.text.y = element_blank())
```

**Plot Notes**

- Francesco Molinari shot 81 in round 4 and 69 in round 3
- Bernd Wiesberger shot 78 in round 4 and 66 in round 2
- Carlos Ortiz shot 82 in round 1 and 71 in round 2 (missed cut)
- Joe Long shot 82 in round 1 and 72 in round 2 (missed cut)
- Hudson Swafford shot 83 in round 2 and 73 in round 1 (missed cut)

___

### Great Shots

Who hit the most great shots?

```{r fold_code = T}
# cor_df = df |>
#   filter(!is.na(great_shots) & !is.na(round_score))
# 
# round(cor(cor_df$great_shots, cor_df$round_score), 3)

df |>
  filter(!is.na(great_shots)) |>
  group_by(player_name) |>
  summarise(great_shots = sum(great_shots)) |>
  slice_max(great_shots, n = 10) |>
  ggplot(aes(reorder(player_name, great_shots), great_shots)) +
  geom_col(fill = custom_green) +
  geom_text(aes(label = great_shots), size = 3.5, hjust = -0.4) +
  coord_flip() +
  labs(x = NULL, y = "Great Shots",
       title = "Who hit the most great shots?",
       subtitle = "2021 Masters Tournament") +
  theme(axis.text.x = element_blank())
```

**Plot Note**: The number of great shots a player hit had a -0.367 correlation coefficient with that player's round score, meaning that as players hit more great shots, their score decreased - just as we'd expect!

___

### Poor Shots

Who hit the most poor shots?

```{r fold_code = T}
# cor_df = df |>
#   filter(!is.na(poor_shots) & !is.na(round_score))
# 
# round(cor(cor_df$poor_shots, cor_df$round_score), 3)

df |>
  filter(!is.na(poor_shots)) |>
  group_by(player_name) |>
  summarise(poor_shots = sum(poor_shots)) |>
  slice_max(poor_shots, n = 10) |>
  ggplot(aes(reorder(player_name, poor_shots), poor_shots)) +
  geom_col(fill = custom_red) +
  geom_text(aes(label = poor_shots), size = 3.5, hjust = -0.4) +
  coord_flip() +
  labs(x = NULL, y = "Poor Shots",
       title = "Who hit the most poor shots?",
       subtitle = "2021 Masters Tournament") +
  theme(axis.text.x = element_blank())
```

**Plot Note**: The number of poor shots a player hit had a 0.592 correlation coefficient with that player's round score, meaning that as players hit more poor shots, their score increased - this is compared to a -0.367 correlation coefficient between great shots and round score, so the number of poor shots a player hit in a round was more impactful to their score than the number of great shots was.

___

### Great and Poor Shots

What is the relationship between great and poor shots?

```{r fold_code = T}
made_missed_cut = df |>
  filter(!is.na(great_shots) & !is.na(poor_shots)) |>
  count(player_name) |>
  mutate(made_cut = ifelse(n == 4, "Made Cut", "Missed Cut"))

# cor_df = df |>
#   filter(!is.na(great_shots) & !is.na(poor_shots)) |>
#   group_by(player_name) |>
#   summarise(great = sum(great_shots),
#             poor = sum(poor_shots),
#             .groups = "drop")
# 
# round(cor(cor_df$great, cor_df$poor), 3)

df |>
  filter(!is.na(great_shots) & !is.na(poor_shots)) |>
  inner_join(made_missed_cut, by = "player_name") |>
  group_by(player_name, made_cut) |>
  summarise(great = sum(great_shots),
            poor = sum(poor_shots),
            .groups = "drop") |>
  ggplot(aes(great, poor)) +
  geom_point(aes(col = made_cut), size = 4) +
  geom_abline(linetype = "dashed", alpha = 0.5) +
  # geom_line(stat = "smooth", formula = y ~ x, method = "lm", linetype = "dashed", alpha = 0.5) +
  scale_color_manual(values = c(custom_green, custom_red)) +
  labs(x = "Great Shots", y = "Poor Shots", col = NULL,
       title = "Scatterplot of Great and Poor Shots",
       subtitle = "2021 Masters Tournament") +
  theme(legend.position = "right") +
  scale_x_continuous(breaks = seq(0, 50, by = 2)) +
  scale_y_continuous(breaks = seq(0, 50, by = 2))
```

**Plot Note**: Over the course of the tournament, there is a positive correlation of 0.583 between the number of great shots a player hit and the number of poor shots a player hit. In the plot, the dashed line represents the points where the number of great shots is equal to the number of poor shots a player hit - ideally, players want to be below this line, indicating they hit more great shots than poor shots.







